---
title: "p8105_hw2_ly2565"
author: "Lin Yang"
date: "10/02/2021"
output: github_document
---

```{r}
library(tidyverse)
library(readxl)
```

# Problem 1

## Read and clean the Mr. Trash Wheel dataset.

```{r}
trash_wheel_df = read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "Mr. Trash Wheel", range = "A2:N408") %>%  
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(sports_balls = round(sports_balls, digits = 0))

trash_wheel_df
```

## Data manipulation on precipitation data for 2018 and 2019.

```{r}
precip_2018_df = 
  read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "2018 Precipitation", range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  mutate(year = 2018) %>% 
  relocate(year)

precip_2019_df = 
  read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "2019 Precipitation", range = "A2:B8") %>% 
  janitor::clean_names() %>% 
  mutate(year = 2019) %>% 
  relocate(year)
```

## Combine precipaitation datasets.

```{r}
comb_precip_df = 
  bind_rows(precip_2018_df, precip_2019_df) %>% 
  mutate(month = month.name[month])
```

* The 2018 precipitation dataset has `r nrow(precip_2018_df)` row and `r ncol(precip_2018_df)` columns with variable names: `r names(precip_2018_df)`. It has precipitation data for all 12 months of 2018. The total precipitation in 2018 is `r sum(pull(precip_2018_df, total))` inches. 
* The 2019 precipitation dataset has `r nrow(precip_2019_df)` row and `r ncol(precip_2019_df)` columns with variable names: `r names(precip_2019_df)`. It has precipitation data for only 6 months of 2019 (Jan-Jun). The total precipitation in 2019 is `r sum(pull(precip_2019_df, total))` inches.
* The median number of sports balls in a dumpster in 2019 is `r median(pull(trash_wheel_df, sports_balls))`.

# Problem 2

## Clean pols_month dataset.

```{r}
pols_month_df = 
  read_csv(file = "data/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, into = c("year", "month", "day")) %>% 
  mutate(
    month = month.name[as.numeric(month)],
    president = prez_gop + prez_dem) %>%
  select(-prez_gop, -prez_dem, -day)
pols_month_df  
```

## Clean snp dataset.

```{r}
snp_df = 
  read_csv(file = "data/snp.csv") %>% 
  mutate(date, date = as.Date(date, format = "%d/%m/%y")) %>% 
  separate(date, into = c("year", "month", "day")) %>% 
  mutate(month = month.name[as.numeric(month)]) %>% 
  select(-day) %>% 
  relocate(year)
snp_df
```

## Clean unemployment dataset.

```{r}
unemploy_df =
  read_csv(file = "data/unemployment.csv") %>% 
  pivot_longer(
    Jan:Dec,
    names_to = "month",
    values_to = "unemploy_pct") %>% 
  janitor::clean_names()
unemploy_df
``` 

## Merge snp into pols, and then merge unemployment.





