---
title: "p8105_hw2_ly2565"
author: "Lin Yang"
date: "10/02/2021"
output: github_document
---

```{r}
library(tidyverse)
library(readxl)
```

# Problem 1

## Read and clean the Mr. Trash Wheel dataset.

```{r}
trash_wheel_df = read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "Mr. Trash Wheel", range = "A2:N408") %>%  
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(sports_balls = round(sports_balls, digits = 0))

trash_wheel_df
```

## Data manipulation on precipitation data for 2018 and 2019.

```{r}
prcp_2018_df = 
  read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "2018 Precipitation", range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  mutate(year = 2018) %>% 
  relocate(year)

prcp_2019_df = 
  read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "2019 Precipitation", range = "A2:B8") %>% 
  janitor::clean_names() %>% 
  mutate(year = 2019) %>% 
  relocate(year)
```

## Combine precipaitation datasets.

```{r}
comb_prcp_df = 
  bind_rows(prcp_2018_df, prcp_2019_df) %>% 
  mutate(month = month.name[month])
comb_prcp_df
```

* The 2018 precipitation dataset has `r nrow(prcp_2018_df)` row and `r ncol(prcp_2018_df)` columns with variable names: `r names(prcp_2018_df)`. It has precipitation data for all 12 months of 2018. The total precipitation in 2018 is `r sum(pull(prcp_2018_df, total))` inches. 
* The 2019 precipitation dataset has `r nrow(prcp_2019_df)` row and `r ncol(prcp_2019_df)` columns with variable names: `r names(prcp_2019_df)`. It has precipitation data for only 6 months of 2019 (Jan-Jun). The total precipitation in 2019 is `r sum(pull(prcp_2019_df, total))` inches.
* The median number of sports balls in a dumpster in 2019 is `r median(pull(trash_wheel_df, sports_balls))`.

# Problem 2

## Clean pols_month dataset.

```{r}
pols_month_df = 
  read_csv(file = "data/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, into = c("year", "month", "day")) %>% 
  mutate(
    month = month.name[as.numeric(month)],
    president = ifelse(prez_dem == 1, "dem", "gop")) %>%
  select(-prez_gop, -prez_dem, -day)
pols_month_df  
```

## Clean snp dataset.

```{r}
snp_df = 
  read_csv(file = "data/snp.csv") %>% 
  janitor::clean_names() %>% 
  separate(date, into = c("month", "day", "year"), convert = TRUE) %>% 
  mutate(
    year = ifelse(year <= 15, year + 2000, year + 1900),
    month = factor(month.name[month], levels = month.name)) %>% 
  select(-day) %>% 
  relocate(year, month) %>% 
  arrange(year, month) %>% 
  mutate(
    year = as.character(year),
    month = as.character(month)
  )
snp_df
```

## Clean unemployment dataset.

```{r}
unemploy_df =
  read_csv(file = "data/unemployment.csv") %>%
  rename_with(~ month.name, month.abb) %>% 
  pivot_longer(
    January:December,
    names_to = "month",
    values_to = "unemploy_percent") %>% 
  janitor::clean_names() %>% 
  mutate(year = as.character(year))
unemploy_df
``` 

## Merge snp into pols, and then merge unemployment.

```{r}
merging_df =
  left_join(pols_month_df, snp_df, by = c("year", "month")) %>% 
  left_join(unemploy_df, by = c("year", "month"))
merging_df
```

* The `pols_month` dataset has `r nrow(pols_month_df)` rows and `r ncol(pols_month_df)` columns with variable names: `r names(pols_month_df)`. The year range is from `r min(pull(pols_month_df, year))` to `r max(pull(pols_month_df, year))`. 
* The `snp` dataset has `r nrow(snp_df)` rows and `r ncol(snp_df)` columns with variable names: `r names(snp_df)`. The year range is from `r min(pull(snp_df, year))` to `r max(pull(snp_df, year))`. 
* The `unemployment` dataset has `r nrow(unemploy_df)` rows and `r ncol(unemploy_df)` columns with variable names: `r names(unemploy_df)`. The year range is from `r min(pull(unemploy_df, year))` to `r max(pull(unemploy_df, year))`.


# Problem 3

## Load and tidy the popular baby name dataset.

```{r}
pop_baby_name_df = 
  read_csv(file = "data/Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    gender = str_to_lower(gender),
    ethnicity = str_to_title(ethnicity),
    childs_first_name = str_to_title(childs_first_name),
    ethnicity = recode(ethnicity,
                       "Black Non Hisp" = "Black Non Hispanic",
                       "Asian And Paci" = "Asian And Pacific Islander",
                       "White Non Hisp" = "White Non Hispanic")
  ) %>% 
  distinct()
pop_baby_name_df
```

## Create a well-structured, reader-friendly table showing the rank in popularity of the name “Olivia”.

```{r}
olivia_rank_df =
  pop_baby_name_df %>% 
  filter(childs_first_name == "Olivia") %>% 
  select(year_of_birth, ethnicity, rank) %>% 
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "rank") %>% 
  knitr::kable()
olivia_rank_df
```

## Create a similar table showing the most popular name among male children.

```{r}
most_pop_male_name_df =
  pop_baby_name_df %>% 
  filter(rank == 1, gender == "male") %>% 
  select(year_of_birth, ethnicity, childs_first_name) %>% 
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "childs_first_name") %>% 
  knitr::kable()
most_pop_male_name_df  
```

## scatter plot
```{r}
pop_baby_name_df %>%
  filter(year_of_birth == 2016, gender == "male", ethnicity == "White Non Hispanic") %>% 
  ggplot(aes(x = rank, y = count)) +
  geom_point()
```

